---
title: "ESS"
author: "Yun Zhao"
date: "26/06/2022"
output: html_document
---

### Load Packages

```{r echo = FALSE}
#install.packages("here")
library(here)
#install.packages("dplyr")
library(dplyr) # data manipulation
#install.packages("tidyr")
library(tidyr) # data manipulation
#install.packages("tidyverse")
library(tidyverse) # data manipulation
#install.packages("leaflet")
library(leaflet) # interactive plotting
#install.packages("sf")
library(sf) # spatial 
#install.packages("naniar")
library(naniar)
library(DDIwR)
library(haven)
#install.packages("janitor")
library(janitor)
#install.packages("dismo")
#library(dismo)
#install.packages("deldir")
library(deldir)
#install.packages("raster")
#install.packages(c("Rcpp", "rgdal", "sp", "terra", "raster"))
install.packages('visibly')
#library(col2hex)
library(visibly)
```

```{r eval= FALSE, echo= FALSE}
#DDIwR::convert(here::here("Data/ESS9MDWe03.1.stata/ESS9MDWe03.1_F1.dta"), to = "Excel")

#ess9multi <- read_dta("Data/ESS9MDWe03.1-20220608.stata/ESS9MDWe03.1_F1.dta") %>%
 # dplyr::select(c(0:10))

#sample(ess9multi, 10)

```

Novembre Genetics
https://raw.githubusercontent.com/NovembreLab/Novembre_etal_2008_misc/master/files/PCA.txt
```{r echo = FALSE}
pca <- read.delim(here::here("Data/Nov2008_PCA-20220622.txt")) 
#samples <- pca %>% dplyr::select(c(1:2))
#samples
nov_colors <- read.delim(here::here("Data/ColorTablePCmap-20220627.txt"), header = F)
nov_colors[nov_colors == "red#lightblue"] <- "red"
novcolors = col2hex(nov_colors[1])

eigs <- read_fwf(here::here("Data/NovPCAeigs-20220627.txt")) %>% 
  row_to_names(., row_number = 1, remove_row = TRUE, remove_rows_above = TRUE) %>%
  clean_names() %>%
  separate(., col = "number_eigvals", into = c("ID","ID2"), sep = " ") %>%
  select(., !matches("ID2"))

predict <- read.csv(here::here("Data/Nov2008_predict-20220627.csv")) %>%  
  rename(predictID = V1, longitude = V2, latitude = V3) %>%
  merge(., samples, by.x = "predictID", by.y = "ID") %>%
  merge(., eigs, by.x = "predictID", by.y = "ID") %>%
  merge(., nov_colors, by.x ="plabels", by.y = "V1", ) %>%
  rename(novcolor = V2) %>%
  sf::st_as_sf(., coords = c("longitude","latitude")) %>%
  st_set_crs(., 4326)

true <- read.csv(here::here("Data/Nov2008_true-20220627.csv")) %>%
  rename(trueID = V1, longitude = V2, latitude = V3) %>%
  merge(., samples, by.x = "trueID", by.y = "ID") %>%
  sf::st_as_sf(., coords = c("longitude","latitude")) %>%
  st_set_crs(., 4326)

predict_sp <- as_Spatial(predict)
#predict_polygons <- voronoi(predict_sp)
```


```{r}
novpal <- colorFactor(nov_colors$V2, nov_colors$V1, ordered = T)

pca_map <- leaflet() %>%
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>% #Esri.WorldGrayCanvas
  #addCircleMarkers(data = true, radius = 5, color = ~novpal(plabels), stroke = T, fillOpacity = 0.1) %>%
  addCircleMarkers(data = predict, radius = 2, color = ~novpal(plabels), stroke = FALSE, fillOpacity = 1) %>%
  addPolygons(data = predict_polygons, fillOpacity = 0, color = ~novpal(plabels), weight = 1, smoothFactor = 0.2)

pca_map
```


```{r}
nov_colors
predict
```

###Load ESS Data 


ess9
https://ess-search.nsd.no/CDW/download
ess9stata
https://adriandusa.eu/post/rename-ess-variables/

```{r echo = FALSE}
ess9_ <- read.csv(here::here("Data/ess9-20220626.csv"))
ess9 <- ess9_ %>%
  dplyr::select(-matches("name"), 
                -matches("essround"),
                -matches("edition"),
                -matches("proddate"),
                -matches("idno"),
                -matches("dweight"),
                -matches("pspwght"),
                -matches("pweight")) %>%
  replace_with_na_all(condition = ~.x == 7) %>%
  group_by(region, na.rm=TRUE) %>%
  summarise(cntry = cntry, across(regunit:impfun, ~ mean(.x, na.rm=TRUE))) %>%
  ungroup() %>%
  distinct()
```

###Load NUTS Data and Merge

nuts 2021, 20M
https://ec.europa.eu/eurostat/web/gisco/geodata/reference-data/administrative-units-statistical-units/nuts

About nuts 
https://ec.europa.eu/eurostat/web/nuts/background
1500 NUTS regions = 92 NUTS 1 + 242 regions NUTS 2 + 1166 regions NUTS 3 

```{r echo = FALSE}
nuts <- st_read(here::here("Data/NUTS_RG_20M_2021_4326-20220626.geojson"))
ess9_nuts <- merge(x = nuts, y = ess9, by.x = "NUTS_ID", by.y = "region") #all.x=TRUE
#ess9_nuts_tidy <- ess9_nuts %>% pivot_longer(cols = ipcrtiv:impfun, names_to = "hvs" )

sample(ess9_nuts, 10)
```

Data Exploration
```{r echo = FALSE, eval = FALSE}
summary(nuts)
colnames(nuts)
sample(nuts$NUTS_ID, 10)
sample(ess9$region, 10)
nuts %>% count(LEVL_CODE)
nrow(ess9)
nrow(nuts)
nrow(ess9_nuts)
sample(ess9_nuts, 10)
colnames(ess9_)    
colnames(ess9) 
```

### Subset by Nuts Level

```{r echo = FALSE}
ess9_nuts1 <- ess9_nuts %>% filter(LEVL_CODE == 1)
ess9_nuts2 <- ess9_nuts %>% filter(LEVL_CODE == 2)
ess9_nuts3 <- ess9_nuts %>% filter(LEVL_CODE == 3)
```

Data Exploration
```{r echo = FALSE, eval = FALSE}
unique(ess9_nuts1$LEVL_CODE)
unique(ess9_nuts2$LEVL_CODE)
unique(ess9_nuts3$LEVL_CODE)
nrow(ess9_nuts1)
nrow(ess9_nuts2)
nrow(ess9_nuts3)
```

### ipcrtiv
```{r echo = FALSE}
pal <- colorNumeric( palette =  "Greens", domain = 1:8 )
pal(c(1,2,3,4,5,6,7,8))

results <- leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron, group = "Positron") %>%
  addPolygons(data = ess9_nuts1, fillColor = ~pal(ipcrtiv), color = "white", weight = 1, fillOpacity = 0.9, smoothFactor = 0.2, group = "NUTS1") %>%
  addPolygons(data = ess9_nuts2, fillColor = ~pal(ipcrtiv), color = "white", weight = 1, fillOpacity = 0.9, smoothFactor = 0.2, group = "NUTS2") %>%
  addPolygons(data = ess9_nuts3, fillColor = ~pal(ipcrtiv), color = "white", weight = 1, fillOpacity = 0.9, smoothFactor = 0.2, group = "NUTS3") %>%
  addLayersControl(
    baseGroups = c("Positron"),
    overlayGroups = c("NUTS1", "NUTS2", "NUTS3"),
    options = layersControlOptions(collapsed = FALSE)
  )

results
```
                                      
### 21 Human Values
```{r echo = FALSE}

pal2 <- colorFactor(topo.colors(5), pca$plabels)

results2 <- leaflet(ess9_nuts) %>%
  addProviderTiles(providers$CartoDB.Positron, group = "Positron") %>%
  
  addPolygons(fillColor = ~pal(ipcrtiv), color = "white", weight = 1, fillOpacity = 0.9, smoothFactor = 0.2, group = "ipcrtiv") %>%
  addPolygons(fillColor = ~pal(imprich), color = "white", weight = 1, fillOpacity = 0.9, smoothFactor = 0.2, group = "imprich") %>%
  addPolygons(fillColor = ~pal(ipeqopt), color = "white", weight = 1, fillOpacity = 0.9, smoothFactor = 0.2, group = "ipeqopt") %>%
  
  addPolygons(fillColor = ~pal(ipshabt), color = "white", weight = 1, fillOpacity = 0.9, smoothFactor = 0.2, group = "ipshabt") %>%
  addPolygons(fillColor = ~pal(impsafe), color = "white", weight = 1, fillOpacity = 0.9, smoothFactor = 0.2, group = "impsafe") %>%
  addPolygons(fillColor = ~pal(impdiff), color = "white", weight = 1, fillOpacity = 0.9, smoothFactor = 0.2, group = "impdiff") %>%
  
  addPolygons(fillColor = ~pal(ipfrule), color = "white", weight = 1, fillOpacity = 0.9, smoothFactor = 0.2, group = "ipfrule") %>%
  addPolygons(fillColor = ~pal(ipudrst), color = "white", weight = 1, fillOpacity = 0.9, smoothFactor = 0.2, group = "ipudrst") %>%
  addPolygons(fillColor = ~pal(ipmodst), color = "white", weight = 1, fillOpacity = 0.9, smoothFactor = 0.2, group = "ipmodst") %>%
  
  addPolygons(fillColor = ~pal(ipgdtim), color = "white", weight = 1, fillOpacity = 0.9, smoothFactor = 0.2, group = "ipgdtim") %>%
  addPolygons(fillColor = ~pal(impfree), color = "white", weight = 1, fillOpacity = 0.9, smoothFactor = 0.2, group = "impfree") %>%
  addPolygons(fillColor = ~pal(iphlppl), color = "white", weight = 1, fillOpacity = 0.9, smoothFactor = 0.2, group = "iphlppl") %>%
  
  addPolygons(fillColor = ~pal(ipsuces), color = "white", weight = 1, fillOpacity = 0.9, smoothFactor = 0.2, group = "ipsuces") %>%
  addPolygons(fillColor = ~pal(ipstrgv), color = "white", weight = 1, fillOpacity = 0.9, smoothFactor = 0.2, group = "ipstrgv") %>%
  addPolygons(fillColor = ~pal(ipadvnt), color = "white", weight = 1, fillOpacity = 0.9, smoothFactor = 0.2, group = "ipadvnt") %>%
  
  addPolygons(fillColor = ~pal(ipbhprp), color = "white", weight = 1, fillOpacity = 0.9, smoothFactor = 0.2, group = "ipbhprp") %>%
  addPolygons(fillColor = ~pal(iprspot), color = "white", weight = 1, fillOpacity = 0.9, smoothFactor = 0.2, group = "iprspot") %>%
  addPolygons(fillColor = ~pal(iplylfr), color = "white", weight = 1, fillOpacity = 0.9, smoothFactor = 0.2, group = "iplylfr") %>%
  
  addPolygons(fillColor = ~pal(impenv), color = "plabels", weight = 1, fillOpacity = 0.9, smoothFactor = 0.2, group = "impenv") %>%
  addPolygons(fillColor = ~pal(imptrad), color = "plabels", weight = 1, fillOpacity = 0.9, smoothFactor = 0.2, group = "imptrad") %>%
  addPolygons(fillColor = ~pal(impfun), color = "plabels", weight = 1, fillOpacity = 0.9, smoothFactor = 0.2, group = "impfun") %>%
  
  addCircleMarkers(data = true, radius = 4, color = ~pal2(plabels), stroke = FALSE, fillOpacity = 1, group = "true")%>%
  addCircleMarkers(data = predict, radius = 4, color = ~pal2(plabels), stroke = FALSE, fillOpacity = 1, group = "predict")%>%
  
  addLayersControl(
    baseGroups = c("Positron"),
    overlayGroups = c("ipcrtiv", "imprich", "ipeqopt", "ipshabt", "ipeqopt", "impsafe", "impdiff",
                      "ipfrule", "ipudrst","ipmodst","ipgdtim","impfree", "iphlppl","ipsuces",
                      "ipstrgv","ipadvnt","ipbhprp","iprspot","iplylfr","impenv", "imptrad", "impfun",
                      "true", "predict"),
    options = layersControlOptions(collapsed = FALSE)
  )

results2
```

### Map
https://stackoverflow.com/questions/39621165/filter-data-points-on-a-map-using-r-leaflet-and-shiny-multiple-criterialibrary(data.table)
```{r echo = FALSE, eval = FALSE}
shinyApp(
  ui = fluidPage(
    sidebarPanel(
      h5("", width = 2),
      checkboxGroupInput(
        inputId = "MatFlag",
        label = h4("NUTs"),
        choices = setNames(
          object = c("NUTS1", "NUTS2", "NUTS3"),
          nm = c("NUTS1", "NUTS2", "NUTS3")
          ),
        ),
        checkboxGroupInput(
          inputId = "AppFlag",
          label = h4("Human Value Scale"),
          choices = setNames(
            object = c(
              "ipcrtiv",
              "imprich",
              "ipeqopt",
              "ipshabt",
              "impsafe",
              "impdiff",
              "ipfrule",
              "ipudrst",
              "ipmodst",
              "ipgdtim",
              "impfree",
              "iphlppl",
              "ipsuces",
              "ipstrgv",
              "ipadvnt",
              "ipbhprp",
              "iprspot",
              "iplylfr",
              "impenv",
              "imptrad",
              "impfun"
            ),
            nm = c(
              "ipcrtiv",
              "imprich",
              "ipeqopt",
              "ipshabt",
              "impsafe",
              "impdiff",
              "ipfrule",
              "ipudrst",
              "ipmodst",
              "ipgdtim",
              "impfree",
              "iphlppl",
              "ipsuces",
              "ipstrgv",
              "ipadvnt",
              "ipbhprp",
              "iprspot",
              "iplylfr",
              "impenv",
              "imptrad",
              "impfun"
            )
          ),
        ),
               position="left"),
  #App mainPanel content and styles
  mainPanel(fluidRow(leafletOutput(outputId="lmap")))
),

  
  server = function(input, output) {
  #Build leaflet map
  lmap <- leaflet(data=ess9_nuts)%>%
    addProviderTiles("Stamen.TonerLite", 
                     options =providerTileOptions(noWrap = TRUE)) #%>%
   fitBounds(~min(-40), ~min(18), ~max(100), ~max(72))
   
  #Filter data
  datFilt <- reactive({
    MatSearch <- paste0(c('xxx',input$MatFlag),collapse = "|")
    MatSearch <- gsub(",","|",MatSearch)
    ess9_nuts_tidy[grepl(MatSearch,"LEVL_CODE") & "hvs" %in% input$AppFlag]
  })
  
  #Add markers based on selected flags
  observe({
    if(nrow(datFilt())==0) {
      print("Nothing selected")
      leafletProxy("lmap") %>%
        clearMarkerClusters()
    } else { #print(paste0("Selected: ", unique(input$InFlags&input$InFlags2)))
      print("Something Selected")
      leafletProxy("lmap", data=datFilt()) %>% 
        clearMarkerClusters() %>% 
        addPolygons( fillColor = ~pal(datFilt), color = "white", weight = 1,
                     fillOpacity = 0.7, smoothFactor = 0.2)
    }
  }) 
  
  output$lmap <- renderLeaflet(lmap)
}


)

```
#let us save the graph
#write_graph(g_new, file_graph, format = "ncol")  
#write_graph(g_tree2, file_graph, format = "ncol") 



#L1 --> unnamed
#L3, is level 2
# 13 jumps, so need 15 levels (bottom and top). j1 - level 2. 
#can take top 10 or 15 in plot tree. put L15 by hand

#changes to igraph?

# L0 disconnected
# ask valentina. 
